r5r: rapid realistic routing on multimodal transport networks with R5 in R

### 2020-09-26 r-package/paper2020/paper2020.Rmd
comment to paper

The package can be installed with the following commands:

@@

# development version
devtools::install_github("ipeaGIT/r5r", subdir = "r-package")
```

@@

Please note that `R` only allocates memory to Java once during a running 
session. If a Java process has been initiated before, either by r5r or any other 
Java-dependent package or function, it might be required to restart the current 
`R` session in order to effectively change the amount of memory allocated.

@@

Building a transport network with r5r is as simple as a one command line:

@@

The function uses the `.pbf` and the `GTFS.zip` files in the directory pointed 
by `data_path` to create a multimodal transport network used for routing by 
R<sup>5</sup>. The resulting `network.dat`, as well as some other files used by 
R<sup>5</sup>, is saved inside the supplied directory.

While the presence of a `.pbf` file is mandatory, since it is required to create 
a street network, the `GTFS.zip` file is only used to calculate public transport 
trips, thus may not be necessary if that's not intended. Multiple GTFS can be 
used simultaneously, in which case all files are merged into a single one 
automatically.

@@

Calculating a travel time matrix with r5r can be done very efficiently with the
`travel_time_matrix()` function. The function takes travel restrictions and
origins/destinations spatial location as inputs and outputs a `data.table` with 
the travel time between all combinations of origins and destinations that can be 
reached with the given restrictions:

```{r, message = FALSE}
# set restrictions
mode <- c("WALK", "TRANSIT")
max_walk_dist <- 1000
max_trip_duration <- 120
departure_datetime <- as.POSIXct("13-03-2019 14:00:00",
                                 format = "%d-%m-%Y %H:%M:%S")

>>>>

The package can be installed with the following command:

@@

```

@@

#> I think this is perhaps unecessary detail for the paper publication.
Please note that `R` only allocates memory to Java once during a running 
session. If a Java process has been initiated before, either by r5r or any other 
Java-dependent package or function, it might be required to restart the current 
`R` session in order to effectively change the amount of memory allocated.

@@

To build a routable transport network with r5r, the user only needs one command
line pointing to `setup_r5()` where the input data is stored:

@@

The function uses the `.pbf` and the `GTFS.zip` files in the directory pointed 
by `data_path` to create a multimodal transport network used for routing by 
R<sup>5</sup>. The resulting `network.dat` as well as some other files used by 
R<sup>5</sup> are saved inside the supplied directory.

The presence of a `.pbf` file with OpenStreetMap data is mandatory because it is
required to create a street network. Meanwhile, a `GTFS.zip` file is only used to
calculate public transport trips, thus may not be necessary if that's not intended.
Multiple GTFS can be used simultaneously, in which case all files are merged into
a single one automatically by R<sup>5</sup>.

@@

Calculating a travel time matrix with r5r can be done very efficiently with the
`travel_time_matrix()` function. The function takes travel parameters and
origins/destinations spatial location as inputs and outputs a `data.table` with 
the travel time between all combinations of origins and destinations that can be 
reached with the given parameters:

```{r, message = FALSE}
departure_datetime <- as.POSIXct("13-03-2019 14:00:00",
                                 format = "%d-%m-%Y %H:%M:%S")

# calculate travel time matrix
ttm <- travel_time_matrix(r5r_core,
                          origins = points,
                          destinations = points,
                          mode = c("WALK", "TRANSIT"),
                          departure_datetime,
                          max_walk_dist = 1000,
                          max_trip_duration = 120)

head(ttm)
```
